Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-24 ribose-paper.RMD
fig1 (fig1.1) finished

Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1). The eligibility criteria were non-smokers and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training (Fig 1). Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

@@

The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2). Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2). Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study, and participants were blinded by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway). A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs. 0g GLU). To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).

@@

The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening. During period I, participants ingested protein and glucose/placebo only. Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs. 0g glucose), and immediately after RT (~0930hrs, 30g vs. 0g glucose). Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g). In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs. 3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose. Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications. The self-chosen diet was repeated on pairwise consecutive days (i.e. on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs. During period III (2200-0700hrs), participants remained in an overnight fasted state. The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously. The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water. During sessions, participants were free to ingest water ad libitum.

@@

Performance tests (Test, Fig 2) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2). Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA). Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5). Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up. The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test. The highest peak torque values were carried forwards to the final analyses. During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day. During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session. Test I at T4 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.

Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2). The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests. Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks. All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions. Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range. Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved. Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range. Attempts with exaggerated hip movement or beneath full extension were not approved.

@@

Muscle biopsies were sampled from m. vastus lateralis using well-established procedures (HAMMARSTRØM20). Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun. After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses. Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4). At each time point, two samples were taken from the same incision. To standardize this procedure, all individual participants had biopsies taken at the same time of day, in an overnight fasted state.

@@

Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.

>>>>

Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth. The eligibility criteria were non-smokers and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training. Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

@@

The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2). Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2). Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study. THe participants were blinded via diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway). A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs. 0g GLU). To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).

@@

The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening. During period I, participants ingested protein and glucose/placebo only. Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs. 0g glucose), and immediately after RT (~0930hrs, 30g vs. 0g glucose). Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g). In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs. 3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose. Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications. The self-chosen diet was repeated on pairwise consecutive days (i.e. on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs. During period III (2200-0700hrs), participants remained in an overnight fasted state. The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously. The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml of water. During sessions, participants were free to ingest water ad libitum.

@@

Strength tests (Fig 2) were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1). Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA). Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5). Isokinetic peak torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up. Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test. The highest peak torque values were carried forwards to the final analyses. During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day. During days 11 and 12, humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session. Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.

Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 1). The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests. Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks. All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions. Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range. Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved. Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range. Attempts with exaggerated hip movement or beneath full extension were not approved.

@@

Muscle biopsies were sampled from m. vastus lateralis using well-established procedures [@hammarstrom.etal_2020]. Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun. After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses. Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4). At each time point, two samples were taken from the same incision. To standardize this procedure, all individual participants had biopsies taken at the same time of day, in an overnight fasted state.

@@

Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study. <!-- Is this necessary to include at all? -->
